---
title: "01_summary_MA Migration"
author: "Gen Perkins"
date: "2023-09-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Summary of Marsh Mispillion Harour DE

```{r library, warning=FALSE, message=FALSE, include = FALSE}
library(leaflet)
library(RColorBrewer)
library(dplyr)
library(lubridate)
library(sf)
library(adehabitatHR)
library(ggplot2)
library(stringr)
library(readxl)
library(dplyr)
library("rnaturalearth")
library("rnaturalearthdata")


#raw_dat <- file.path("data", "movebank_locations_20230407")
#raw_dat <- file.path("data", "movebank_locations_20230724")
raw_dat <- file.path("data", "movebank_locations_20231219")

filesoi <- list.files(raw_dat)

# data_set3 : spring migration 
key = "Mispillion Harbor, DE"

filesoi <- list.files(raw_dat, pattern = key)
filesoi_ref <- filesoi[1]
filesoi <- filesoi[2]

# read in reference data 
brep <- read.csv(file.path(raw_dat, filesoi_ref))
brep <- brep %>%
#  dplyr::select(c(tag.id , "animal.id" ,deploy.on.date,
#                   tag.manufacturer.name, tag.mass,   tag.model, tag.serial.no))%>%
mutate(tag.local.identifier = tag.id,
  animal.marker.id = animal.id, 
  animal.id = str_c("MISP_", tag.id, "_2023")) %>% 
  mutate(animal.marker.id = str_sub(animal.marker.id, -9, -1), 
         tag.model = "Lotek PinPoint GPS-Argos 75",
         tag.manufacturer.name = "Lotek") 


#no.tags.ref <- unique(brep$tag.local.identifier)
#no.ids.ref <- unique(brep$animal.id)


#filesoi <- list.files(raw_dat, pattern = key)

btemp <- read.csv(file.path(raw_dat, filesoi))
bout <- btemp %>%
    dplyr::select(visible, timestamp, location.long, location.lat,
                 gps.fix.type.raw, lotek.crc.status,
                  sensor.type, individual.local.identifier,
                   height.above.ellipsoid, import.marked.outlier,
                  argos.lc, algorithm.marked.outlier,argos.altitude,
                  height.above.ellipsoid, tag.local.identifier) %>%
  mutate(date_time = ymd_hms(timestamp))%>%
  mutate(year = year(date_time )) %>%
  mutate(month = month(date_time ),
         day = day(date_time),
         hour = hour(date_time),
         minute = minute(date_time))


all_dat <- left_join(bout, brep )%>% 
  dplyr::select(-individual.local.identifier, 
         -deploy.off.date, - animal.taxon, -tag.local.identifier ) %>%
  mutate(argos.lc = as.character(argos.lc))%>%
  mutate(deploy.on.latitude = 39.20075, 
       deploy.on.longitude = -75.0255 , 
       study.site = "Thomp")%>%
  mutate(tag.id = as.character(tag.id))
  

head(all_dat)

all_dat <- all_dat %>%
  mutate(id = seq(1, length(all_dat$visible), 1))


##############################
# durations

# Duration between pings/ 
bdd <- all_dat |> 
  mutate(ddate = ymd_hms(date_time)) |> 
  arrange(tag.id, ddate)

bdd_dur <- bdd  |> 
  group_by(tag.id) |> 
  mutate(diff = difftime(ddate, lag(ddate),  units = c("hours")), 
         diff = as.numeric(diff))%>%
  dplyr::filter(diff >0)


## Calculate distance between points and bearing

bdd_det <- bdd_dur  |> 
  #filter(tag.id == 230318) |> 
  group_by(tag.id) |> 
  mutate(location.long_prior = lag(location.long, 1L),
         location.lat_prior = lag(location.lat, 1L)) %>%
  rowwise() %>%
  mutate(gcd_m = distHaversine(c(location.long_prior,location.lat_prior), c(location.long, location.lat)),
         bearing = bearing(c(location.long_prior,location.lat_prior), c(location.long, location.lat)),
         speed_mhr = round((gcd_m/diff)/1000,1))%>%
  ungroup()

# determine the location of direction 

head(bdd_det)


# stop over = within 25 km of previous point? 

bt <- bdd_det |> 
  dplyr::select(id, tag.id, ddate, day, month, year,location.long, location.lat,location.lat_prior, diff, gcd_m,bearing, speed_mhr) %>%
  group_by(tag.id)%>%
  mutate(stopover = ifelse( gcd_m <= 25000, "stop-over", "migration")) %>%
  mutate(breeding = case_when(
    stopover == "stop-over" & month %in% c(6,7,8) & location.lat > 60 ~ "breeding", 
    .default = "NA")) %>%
 mutate(direction = case_when(
    location.lat >= location.lat_prior ~ "northward", 
    location.lat <= location.lat_prior~ "southward",
    .default = "NA"))%>%
  ungroup()


idll <- bt|> 
  dplyr::select("location.long", "location.lat", id)


clean_sf <- st_as_sf(bt, coords = c("location.long", "location.lat"), crs = st_crs(4326))
st_write(clean_sf, file.path("output", "pt_misoillion_20240123.gpkg"), append = F)

# 

## Post manual edits

manual_edits <- file.path("output", "manual_edited_complete")
milsedit <- read.csv(file.path( manual_edits, "pt_mils_mig_all.csv")) %>%
  left_join(idll) %>%
   mutate(direction = case_when(
           breeding == "wintering"  ~ NA, 
           stopover == "breeding" ~ NA,
          .default = as.character(direction)))%>% 
  dplyr::filter(!is.na(location.lat)) %>%
  dplyr::select(-tag.id, -ddate, -day, -month, -year, -"location.long", -"location.lat", -fid)
 

head(milsedit)
milout <- left_join(all_dat, milsedit, by = "id")

# #save out file
clean_save = milout %>% mutate(proj = "Mispillion")
saveRDS(clean_save, file = file.path("output", "rekn_mispillion_20240123.rds"))


clean_sf <- st_as_sf(clean_save, coords = c("location.long", "location.lat"), crs = st_crs(4326))
st_write(clean_sf, file.path(manual_edits, "rekn_mils_mig_20240123.gpkg"), append = F)


```

This dataset has `r length(no.tags)` tags (tag.local.identifier) and `r length(no.ids)` (individual.local.identifier). 

```{r, echo = FALSE}
knitr::kable(ids)

```

```{r, echo=FALSE}

clean <- bout %>%
  filter(visible == "true")%>%
  filter(!is.na(location.long)) %>%
  filter(!is.na(location.lat))


cleanargos <- clean %>% 
  dplyr::filter(argos.lc != "Z")%>%
  dplyr::filter(argos.lc != "A")%>% 
 dplyr::filter(argos.lc != "B")

#length(cleanargos$visible)

clean = cleanargos

errrors_tble <- clean %>%
  group_by(tag.local.identifier, year, visible) %>%
  summarise(count = n())


# #save out file
clean_save = clean %>% mutate(proj = "Mispiloin_harbour")
saveRDS(clean_save, file = file.path("output", "rekn_mispiloin_20231102.rds"))



```


# Raw data

The number of raw records as of September 2023 is `r length(bout$tag.local.identifier)`.  The raw date range varies from 
`r min(range(bout$year))` to `r max(range(bout$year))`.


If we remove all the algorithm marked outliers (lotek = error, argos accuracy value = A, B, Z). We reduce the number of observations in the data set from `r length(bout$tag.local.identifier)` to `r length(clean$tag.local.identifier)`.
Argos accuracy codes are: 

- Class 0: over 1500 m radius (not currently filtered)

- Class A: Argos no accuracy estimation (3 messages received)
- Class B: Argos no accuracy estimation (1 or 2 messages received)
- Class Z: Argos Invalid Location


The cleaned date range varies from 
`r min(range(clean$year))` to `r max(range(clean$year))`.


# GPS vs argos


The data split between gps and argos data records. Large proportion are argos doppler shift.   These represent two different methods of calculating the locations and all types can be used. 

```{r, echo = FALSE, message=FALSE}
tag_type <- clean %>%
  dplyr::select(tag.local.identifier, sensor.type) %>%
  group_by(tag.local.identifier, sensor.type)%>%
  summarise(n = n())

#tag_type

tag_type_date <- clean %>%
  dplyr::select(tag.local.identifier, sensor.type, year, month) %>%
   group_by(tag.local.identifier, sensor.type, year,month)%>%
  summarise(n = n())

#tag_type_date

#p_alldat <- ggplot(clean, aes(year, fill = sensor.type))+
#  geom_bar(position = "dodge") 

p_alldat <- ggplot(clean, aes(year, fill = sensor.type))+
  geom_bar(position = "dodge") +
  xlim(2021, 2024)+
  facet_wrap(~tag.local.identifier)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

This seems odd as raw data shows shift from gps Lotek tag to argos-dopler-shift? 

```{r}
p_alldat 
```


# Duration of Tag summary

```{r, echo = FALSE}


table_max <- clean %>% 
  dplyr::select(tag.local.identifier, date_time) %>%
 # group_by(tag.local.identifier)%>% 
  slice_max(date_time, by = tag.local.identifier)
colnames(table_max)<- c("tag.local.identifier", "max")

table_min <- clean %>% 
  dplyr::select(tag.local.identifier, date_time) %>%
 # group_by(tag.local.identifier)%>% 
  slice_min(date_time, by = tag.local.identifier)

colnames(table_min)<- c("tag.local.identifier", "min")

dur <- left_join(table_max, table_min,by = join_by(tag.local.identifier)) %>%
  distinct() %>%
  dplyr::mutate(duration = max- min)%>%
  mutate(dur_days = round(as.numeric(duration),1))%>%
  mutate(year = year(min))


dur_data = dur


dur_plot <- ggplot(dur, aes(y=factor(tag.local.identifier))) +
  geom_segment(aes(x=min, xend=max, y=factor(tag.local.identifier),    yend=factor(tag.local.identifier)), size=1)+
  xlab("Date") + ylab("Tag") 

dur_plot

dur_hist <- ggplot(dur, aes(dur_days))+
  geom_histogram(binwidth = 5) +
  xlab("duration (days)") 

dur_hist

p_duration <- ggplot(clean, aes(factor(month), fill = factor(year)))+
  geom_bar(position = "dodge") +
  #xlim(1,12)+
  facet_wrap(~tag.local.identifier)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

p_duration

```


# Number of active tags 

At the last data download, August 1st, a number of tags were continuing to record locations. THese represent active tags where more details can be gained. Inactive tags are those which have stopped transmitting. 

```{r , echo=FALSE}
#head(dur_data)

library(dplyr)
ddur <- dur_data %>%
  dplyr::mutate(max_month = month(max),
         max_year = year(max))
ddur <- ddur %>%
  dplyr::mutate(active = ifelse(max_month == 9 & max_year == 2023, "y", "n"))

active = ddur %>% filter(active == "y")
inactive = ddur %>% filter(active == "n")

```


Of the `r length(ddur$tag.local.identifier)` tags, `r length(active$tag.local.identifier)` were active, and `r length(inactive$tag.local.identifier)` were inactive.



# Geographic distributon of tags

```{r, echo=FALSE}
rf_sf <- sf::st_as_sf(clean, coords = c("location.long","location.lat"), crs = 4326, agr = "constant")

world <- ne_countries(scale = "medium", returnclass = "sf")

Americas <- world %>% dplyr::filter(region_un == "Americas")
#Americas <- world %>% dplyr::filter(continent == "North America")

# entire north America 
global <- ggplot(data = Americas) +
  geom_sf(color = "grey") +
  geom_sf(data = rf_sf, size = 1, colour = "dark blue") +
  facet_wrap(~tag.local.identifier)+
  # geom_point(ru, aes(x = lng, y = lat), size = 4) +
  xlab("Longitude") + ylab("Latitude") +
  coord_sf(xlim = c(-110, -30), ylim = c(-10, 80), expand = FALSE)+
  theme_bw()+
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank())

global

# 
# jpeg(file.path(out.plots,"rose_density.jpg"), width = 30, height = 30,units = "cm", res = 210)
# # 2. Create the plot
# global 
```



# Detailed geographic locations vs sub populations

We can estimate the subpopulation level for those individuals with longer duration of tag timing?
Using the estimated ranges
- South East US
- Northern South America
- Western Gulf and western Pacific. 


```{r, echo = FALSE}
# 
# all <- alldat %>% dplyr:: select(individual.local.identifier.x,dur_days, year, active, dur_class,gcd, gcd_class)%>%
#   rename("individual.local.identifier"  = individual.local.identifier.x )
#    
# arf_sf <- left_join(rf_sf, all) 

a <- sort(unique(clean$tag.local.identifier ))



baffin <- c( 236449)

kw <- c(236448)

pvic <- c(236453)
         
cari <- c(236444)

sth_am <- c(236445,  236450, 236451)
            
hudson <- c(236452)
  
odd <- c( 236446, 236447) 



# breeding local 

breedsf <- rf_sf %>% 
  mutate(breed_class = case_when(
     tag.local.identifier %in%  baffin ~ "baffin",
     tag.local.identifier %in%  hudson ~ "hudson_bay",
     tag.local.identifier %in% sth_am ~ "sth_america",
     tag.local.identifier %in%  cari ~ "caribbean",
     tag.local.identifier %in%  kw ~ "king_william",
     tag.local.identifier %in%  pvic ~ "princess_vic",
      tag.local.identifier %in%  odd ~ "unknown",
      .default = NA
))


# entire north America 
breed_loc <- ggplot(data = Americas) +
  geom_sf(color = "grey") +
  geom_sf(data = breedsf, size = 1, colour = "dark blue") +
  facet_wrap(~breed_class)+
  # geom_point(ru, aes(x = lng, y = lat), size = 4) +
  xlab("Longitude") + ylab("Latitude") +
 coord_sf(xlim = c(-110, -30), ylim = c(-10, 80), expand = FALSE)+
  theme_bw()+
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank())

breed_loc

```




