---
title: "01_Summary_OceanWinds"
author: "Gen Perkins"
date: "2023-09-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary of Ocean Winds

```{r library, warning=FALSE, message=FALSE, include = FALSE}
library(leaflet)
library(RColorBrewer)
library(dplyr)
library(lubridate)
library(sf)
library(adehabitatHR)
library(ggplot2)
library(stringr)
library(readxl)
library(dplyr)
library("rnaturalearth")
library("rnaturalearthdata")


#raw_dat <- file.path("data", "movebank_locations_20230407")
#raw_dat <- file.path("data", "movebank_locations_20230724")
raw_dat <- file.path("data", "movebank_locations_20231219")

filesoi <- list.files(raw_dat)

# data_set3 : spring migration 
key = "Red Knot Fall Migration"

filesoi <- list.files(raw_dat, pattern = key)
filesoi_ref <- filesoi[1]
filesoi <- filesoi[2]



# read in reference data 
brep <- read.csv(file.path(raw_dat, filesoi_ref))
brep <- brep %>%
    dplyr::select(c(tag.id , "animal.id" ,deploy.on.date, "animal.marker.id","animal.ring.id", "tag.serial.no", animal.life.stage, deploy.on.measurements,  animal.mass))%>%
mutate(tag.local.identifier = tag.id)

no.tags.ref <- unique(brep$tag.local.identifier)
no.ids.ref <- unique(brep$animal.id)

# number of unique tags = no of ids 


# read in track data 

btemp <- read.csv(file.path(raw_dat, filesoi))
bout <- btemp %>%
    dplyr::select(visible, timestamp, location.long, location.lat,
                  argos.altitude, import.marked.outlier, 
                  gps.fix.type.raw, lotek.crc.status,lotek.crc.status.text, 
                  sensor.type, individual.local.identifier,
                  argos.lc, algorithm.marked.outlier,
                  tag.local.identifier)


# calculate time differences
bout <- bout  %>%
  mutate(date_time = ymd_hms(timestamp))%>%
  mutate(year = year(date_time )) %>%
  mutate(month = month(date_time ),
         day = day(date_time),
         hour = hour(date_time),
         minute = minute(date_time))

no.tags <- unique(bout$tag.local.identifier)
no.ids <- unique(bout$individual.local.identifier)
  

ids <- bout %>%
  dplyr::select(tag.local.identifier, individual.local.identifier) %>%
  group_by(tag.local.identifier, individual.local.identifier)%>%
  summarise(n = n())
  

ymin = min(range(bout$year))
ymax = max(range(bout$year))

###########################

brep 
bout

all_dat <- left_join(bout, brep )

head(all_dat)


# #save out file
# #save out file
clean_save = all_dat %>% mutate(proj = "OceanWinds")
saveRDS(clean_save, file = file.path("output", "rekn_ocean_20231219.rds"))


```

This dataset has `r length(no.tags)` tags (tag.local.identifier) and `r length(no.ids)` (individual.local.identifier). 


For the reference data there are `r length(no.tags.ref)` tags and `r length(no.ids.ref)` (individual.local.identifier). 


```{r, echo = FALSE}
knitr::kable(ids)

```

```{r, echo=FALSE}

clean <- bout %>%
  filter(visible == "true")%>%
  filter(!is.na(location.long)) %>%
  filter(!is.na(location.lat)) 

#length(clean$visible)

cleanargos <- clean %>% 
  dplyr::filter(argos.lc != "Z")%>%
  dplyr::filter(argos.lc != "A")%>% 
 dplyr::filter(argos.lc != "B")


#length(cleanargos$visible)

clean = cleanargos

# #save out file
clean_save = clean %>% mutate(proj = "OceanWinds")
saveRDS(clean_save, file = file.path("output", "rekn_ocean_20231102.rds"))


```


# Raw data

The number of records as of September 2023 is `r length(bout$tag.local.identifier)`.  The raw date range varies from 
`r min(range(bout$year))` to `r max(range(bout$year))`. 

If we remove all the algorithm marked outliers (lotek = error, argos accuracy value = A, B, Z). We reduce the number of observations in the data set from `r length(bout$tag.local.identifier)` to `r length(clean$tag.local.identifier)`.

- Class A: Argos no accuracy estimation (3 messages received)
- Class B: Argos no accuracy estimation (1 or 2 messages received)
- Class Z: Argos Invalid Location

The cleaned date range varies from 
`r min(range(clean$year))` to `r max(range(clean$year))`.


# GPS vs argos

The data split between gps and argos data records. 

```{r, echo = FALSE, message=FALSE}
tag_type <- clean %>%
  dplyr::select(tag.local.identifier, sensor.type) %>%
  group_by(tag.local.identifier, sensor.type)%>%
  summarise(n = n())

#tag_type

tag_type_date <- clean %>%
  dplyr::select(tag.local.identifier, sensor.type, year, month) %>%
   group_by(tag.local.identifier, sensor.type, year,month)%>%
  summarise(n = n())

#tag_type_date

p_alldat_all <- ggplot(clean, aes(year, fill = sensor.type))+
  geom_bar(position = "dodge") +
    xlim(2020, 2023)

p_alldat <- ggplot(clean, aes(year, fill = sensor.type))+
  geom_bar(position = "dodge") +
  xlim(2020, 2023)+
  facet_wrap(~tag.local.identifier, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```



```{r,echo=FALSE}
p_alldat_all
p_alldat 

```


# Duration of Tag summary

```{r, echo = FALSE}
# 
# p_duration <- ggplot(clean, aes(factor(month), fill = factor(year)))+
#   geom_bar(position = "dodge") +
#   #xlim(1,12)+
#   facet_wrap(~tag.local.identifier)+
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
# 
# p_duration

table_max <- clean %>% 
  dplyr::select(individual.local.identifier, date_time) %>%
 # group_by(tag.local.identifier)%>% 
  slice_max(date_time, by = individual.local.identifier)%>%
  unique()
colnames(table_max)<- c("individual.local.identifier", "max")

table_min <- clean %>% 
  dplyr::select(individual.local.identifier, date_time) %>%
 # group_by(tag.local.identifier)%>% 
  slice_min(date_time, by = individual.local.identifier) %>%
  unique()
colnames(table_min)<- c("individual.local.identifier", "min")

dur <- left_join(table_max, table_min,by = join_by(individual.local.identifier)) %>%
  distinct() %>%
  dplyr::mutate(duration = max- min)%>%
  mutate(dur_days = round(as.numeric(duration),1))%>%
  mutate(year = year(min))

dur_data = dur

 
dur_plot <- ggplot(dur, aes(y=factor(individual.local.identifier))) +
  geom_segment(aes(x=min, xend=max, y=factor(individual.local.identifier),    yend=factor(individual.local.identifier)), size=1)+
  xlab("Date") + ylab("Tag") 

#dur_plot

dur_hist <- ggplot(dur, aes(dur_days))+
  geom_histogram() +
  xlab("duration (days)") 

#dur_hist

p_duration <- ggplot(clean, aes(factor(month), fill = factor(year)))+
  geom_bar(position = "dodge") +
  #xlim(1,12)+
  facet_wrap(~individual.local.identifier)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

#p_duration



```

```{r, echo=FALSE}
dur_plot
dur_hist
p_duration

```

# Banding locations

The banding locations were largely in Delaware Bay and coastal New Jersey (included stone harbour (STOH), Avalon (AVAL), North Brigatine (NBRIG) and ust south of stone harbour (TWOM). 
This table shows the number of individuals by banding location and year.


```{r, echo = FALSE}
#head(dur)

loc <- dur %>%
  dplyr::select(c(individual.local.identifier , dur_days))%>%
  mutate(dloc = stringr::str_sub(individual.local.identifier, 1, 4)) %>%
  mutate(dyr = stringr::str_sub(individual.local.identifier,13, 17)) %>%
  mutate(dyr = gsub("_", "", dyr))

dep_summary <- loc %>%
  dplyr::group_by(dloc, dyr)%>%
  dplyr::summarise( dloc_n = n())

#dep_summary

```

```{r, echo = FALSE}
knitr::kable(dep_summary)
```


# Number of active tags 

At the last data download all tags in this project were finished transmitting.









# Geographic distributon of tags

```{r, echo=FALSE}
rf_sf <- sf::st_as_sf(clean, coords = c("location.long","location.lat"), crs = 4326, agr = "constant")

world <- ne_countries(scale = "medium", returnclass = "sf")

Americas <- world %>% dplyr::filter(region_un == "Americas")
#Americas <- world %>% dplyr::filter(continent == "North America")

# entire north America 
global <- ggplot(data = Americas) +
  geom_sf(color = "grey") +
  geom_sf(data = rf_sf, size = 1, colour = "dark blue") +
  #facet_wrap(~tag.local.identifier)+
  # geom_point(ru, aes(x = lng, y = lat), size = 4) +
  xlab("Longitude") + ylab("Latitude") +
  coord_sf(xlim = c(-110, -30), ylim = c(-10, 80), expand = FALSE)+
  theme_bw()+
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank())

global

# 
# jpeg(file.path(out.plots,"rose_density.jpg"), width = 30, height = 30,units = "cm", res = 210)
# # 2. Create the plot
# global 
```

Comments 
- A few individuals traveled to the Northern South America (NBRIG_233919, NBRIG_233926, NBRIG_233928). 

```{r, echo = FALSE}

a <- sort(unique(clean$individual.local.identifier ))


# sth_am
sth <- c( "NBRI_233926_2022", "NBRI_233928_2022" , "NBRI_233919_2022")
 

breedsf <- rf_sf %>% 
  mutate(breed_class = case_when(
     individual.local.identifier %in%  sth ~ "south_america",
      .default = "USA"
))


# entire north America 
breed_loc <- ggplot(data = Americas) +
  geom_sf(color = "grey") +
  geom_sf(data = breedsf, size = 1, colour = "dark blue") +
  facet_wrap(~breed_class)+
  # geom_point(ru, aes(x = lng, y = lat), size = 4) +
  xlab("Longitude") + ylab("Latitude") +
  coord_sf(xlim = c(-110, -30), ylim = c(-10, 80), expand = FALSE)+
  theme_bw()+
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank())

breed_loc


```


Questions? 

- Reference data shows 61 individuals, only 45 with tracks and only 43 with tracks with sufficient accuracy (ie argos location estimates of 1500m or less). 

- This data set is short duration tags is all less than 30 days

