---
title: "01_Summary_Atlantic_shores"
author: "Gen Perkins"
date: "2023-09-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## Summary of Atlantic Shores

```{r library, warning=FALSE, message=FALSE, include = FALSE}

library(leaflet)
library(RColorBrewer)
library(dplyr)
library(lubridate)
library(sf)
library(adehabitatHR)
library(ggplot2)
library(stringr)
library(readxl)
library(dplyr)
library("rnaturalearth")
library("rnaturalearthdata")


#raw_dat <- file.path("data", "movebank_locations_20230724")
#raw_dat <- file.path("data", "movebank_locations_20230918")
raw_dat <- file.path("data", "movebank_locations_20231219")
filesoi <- list.files(raw_dat)

# data_set3 : spring migration 
key = "Atlantic"

filesoi <- list.files(raw_dat, pattern = key)
filesoi_ref <- filesoi[1]
filesoi <- filesoi[2]

# read in reference data 
brep <- read.csv(file.path(raw_dat, filesoi_ref))
 brep <- brep %>%
#   dplyr::select(c(tag.id , "animal.id" ,deploy.on.date, "animal.marker.id","animal.ring.id", "tag.serial.no", animal.life.stage, deploy.on.measurements,  animal.mass, 
#                   tag.serial.no,tag.manufacturer.name, tag.model))%>%
 mutate(tag.local.identifier = tag.id)

# 
# no.tags.ref <- unique(brep$tag.local.identifier)
# no.ids.ref <- unique(brep$animal.id)
# 
# 
# ids <- brep %>% 
#   dplyr::select(tag.local.identifier, animal.id) %>%
#   distinct()

# based on the reference data there is 
# 107 unique tag ids (numeric) and 77 unique animal ids *ie NBRI_233781_2022


# read in track data

btemp <- read.csv(file.path(raw_dat, filesoi))

bout <- btemp %>%
    dplyr::select(visible, timestamp, location.long, location.lat,
                  algorithm.marked.outlier, argos.altitude, import.marked.outlier, 
                  gps.fix.type.raw, lotek.crc.status,
                  sensor.type, individual.local.identifier,
                  argos.lc, 
                  tag.local.identifier)%>%
  mutate(date_time = ymd_hms(timestamp))


# calculate time differences
#bout <- bout %>%
#  mutate(date_time = ymd_hms(timestamp))%>%
#   mutate(year = year(date_time )) %>%
#   mutate(month = month(date_time ),
#          day = day(date_time),
#          hour = hour(date_time),
#          minute = minute(date_time))
# 
# no.tags <- unique(bout$tag.local.identifier)
# no.ids <- unique(bout$individual.local.identifier)
#   

# which ids are missing tracking data? 

# based on the tracks data there are 57 ids and animal tag nos. 
# all tracks have a animal tag id (ie: )


#brep #107
#bout

all_dat <- left_join(bout, brep )

head(all_dat)#2089


all_dat <- all_dat %>% 
  dplyr::select(-individual.local.identifier, -tag.local.identifier, -animal.taxon, 
         -deploy.off.date, -algorithm.marked.outlier, -import.marked.outlier) %>%
  mutate(tag.id = as.character(tag.id)) %>%
  filter() %>% 
  dplyr::mutate(lotek.crc.status = case_when(
    is.na(lotek.crc.status) ~ "",
    TRUE ~ as.character(lotek.crc.status)))%>%
  dplyr::filter(lotek.crc.status != "E")%>%
  dplyr::mutate(argos.lc = case_when(
    is.na(argos.lc) ~ "",
    TRUE ~ as.character(argos.lc)))%>%
  dplyr::filter(argos.lc != "Z")

unique(all_dat$animal.id)


library(stringr)

all_dat <- all_dat %>% 
  mutate(study.site1 = substr( animal.id, start = 1, stop = 4)) %>%
  dplyr::mutate(deploy.on.latitude = case_when(
    study.site1 == "NBRI" ~ 39.436,
    study.site1 == "Thom" ~ 39.20075,
    study.site1 == "Norb" ~ 39.20075)) %>%
  dplyr::mutate(deploy.on.longitude = case_when(
    study.site1 == "NBRI" ~ -74.346,
    study.site1 == "Thom" ~ -75.0255,
    study.site1 == "Norb" ~ -75.0255))%>%
  mutate(study.site = study.site1) %>%
  dplyr::select(-study.site1)
 
 
# 
# 
# # Nbrig latitutde/ longitide 
# Noth Brigatine 
# -74.346 39.436
# 
# 
# # Thom = thompson beach 
# -75.0255 , 39.20075
# 
# # norb 
#    -75.0255 , 39.20075


# #save out file
# #save out file
clean_save = all_dat %>% mutate(proj = "atlantic")
saveRDS(clean_save, file = file.path("output", "rekn_atlantic_20240101.rds"))



###########################################################################





ids.track <- bout %>% 
  dplyr::select(tag.local.identifier, individual.local.identifier) %>%
  distinct()

fullid <- full_join(ids, ids.track)


ids <- bout %>%
  dplyr::select(tag.local.identifier, individual.local.identifier) %>%
  group_by(tag.local.identifier, individual.local.identifier)%>%
  summarise(n = n())
  

ymin = min(range(bout$year))
ymax = max(range(bout$year))
```

This dataset has some discrepancies within the reference data and the track data. The reference data show 107 tags (based on tag-id number ie. 204561). Of these I found deployment information on 78 tags, which also includes 16 tags which are duplicated in USFW study. Tracks for these 16 duplicated tags are within the atlantic tags track set but appear less updated (unsure why?). 

A review of the track data shows records for `r length(no.tags)` tags (tag.local.identifier) and `r length(no.ids)` (individual.local.identifier). 

This indicates that some tag ids listed in the reference data may not yet have been deployed (i.e 29 tags with no reference data and no deployment data)


```{r, echo = FALSE}
knitr::kable(ids)

```


```{r, echo=FALSE}

errors <- ggplot(bout, aes(year, fill = visible))+
    geom_bar(position = "dodge")#+
    #xlim(2021,2024)#+
  #facet_wrap(~tag.local.identifier)

clean <- bout %>%
  filter(visible == "true")%>%
  filter(!is.na(location.long)) %>%
  filter(!is.na(location.lat)) %>%
  filter(year <2024)


clean <- clean %>% 
  dplyr::filter(argos.lc != "Z")%>%
  dplyr::filter(argos.lc != "A")%>% 
 dplyr::filter(argos.lc != "B")



  
errors <- ggplot(clean, aes(year, fill = visible))+
    geom_bar(position = "dodge")
# errrors_tble <- bout %>%
#   group_by(tag.local.identifier, year, visible) %>%
#   summarise(count = n())



# #save out file
clean_save = clean %>% mutate(proj = "atlantic")
saveRDS(clean_save, file = file.path("output", "rekn_atlantic_20231102.rds"))




```


# Raw data

The number of records as of early June 2023 is `r length(bout$tag.local.identifier)`.  The raw date range varies from 
`r min(range(bout$year))` to `r max(range(bout$year))`. 

```{r, echo = FALSE}
errors
```



If we remove all the algorithm marked outliers (lotek = error, argos accuracy value = A, B, Z). We reduce the number of observations in the data set from `r length(bout$tag.local.identifier)` to `r length(clean$tag.local.identifier)`.
Argos accuracy codes are: 

- Class 0: over 1500 m radius (not currently filtered)

- Class A: Argos no accuracy estimation (3 messages received)
- Class B: Argos no accuracy estimation (1 or 2 messages received)
- Class Z: Argos Invalid Location


The cleaned date range varies from 
`r min(range(clean$year))` to `r max(range(clean$year))`.

# GPS vs argos

The data set contains both gps and argos dopler shift location estimates. These represent two different methods of calculating the locations and all types can be used. 

```{r, echo = FALSE, message=FALSE}
tag_type <- clean %>%
  dplyr::select(tag.local.identifier, sensor.type) %>%
  group_by(tag.local.identifier, sensor.type)%>%
  summarise(n = n())

#tag_type

tag_type_date <- clean %>%
  dplyr::select(tag.local.identifier, sensor.type, year, month) %>%
   group_by(tag.local.identifier, sensor.type, year,month)%>%
  summarise(n = n())

#tag_type_date

p_alldat <- ggplot(clean, aes(year, fill = sensor.type))+
  geom_bar(position = "dodge") 

```

```{r, echo= FALSE}
p_alldat 

```

```{r, echo = FALSE}
# linking the tags to reference data

# read in reference data 
brep <- read.csv(file.path(raw_dat, filesoi_ref))
brep <- brep %>%
  dplyr::select(c(tag.id , "animal.id" ,deploy.on.date, "animal.marker.id","animal.ring.id", "tag.serial.no"))%>%
mutate(tag.local.identifier = tag.id)


alltrack <- clean %>% 
  left_join(brep) %>% 
  dplyr::select(-tag.id, -animal.id, - visible, -timestamp, -algorithm.marked.outlier)%>%
 distinct()

```


# Duration of Tag summary

```{r, echo = FALSE}
# 
# p_duration <- ggplot(clean, aes(factor(month), fill = factor(year)))+
#   geom_bar(position = "dodge") +
#   #xlim(1,12)+
#   #facet_wrap(~tag.local.identifier)+
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
# 
# p_duration

#alltrack

table_max <- alltrack %>% 
  dplyr::select(tag.local.identifier, date_time) %>%
 # group_by(tag.local.identifier)%>% 
  slice_max(date_time, by = tag.local.identifier)
colnames(table_max)<- c("tag.local.identifier", "max")

table_min <- alltrack %>% 
  dplyr::select(tag.local.identifier, date_time) %>%
 # group_by(tag.local.identifier)%>% 
  slice_min(date_time, by = tag.local.identifier)

colnames(table_min)<- c("tag.local.identifier", "min")

dur <- left_join(table_max, table_min,by = join_by(tag.local.identifier)) %>%
  distinct() %>%
  dplyr::mutate(duration = max- min)%>%
  mutate(dur_min = round(as.numeric(duration)/60,1))%>%
  mutate(dur_hrs = round(as.numeric(dur_min)/60,1))%>%
  mutate(dur_days = round( dur_hrs/24,1))%>%
  mutate(year = year(min))

dur_data = dur


dur_plot <- ggplot(dur, aes(y=factor(tag.local.identifier))) +
  geom_segment(aes(x=min, xend=max, y=factor(tag.local.identifier), yend=factor(tag.local.identifier)), size=1)+
  xlab("Date") + ylab("Tag") 

dur_plot

dur_hist <- ggplot(dur, aes(x= dur_days, colour = as.factor(year)))+
  geom_histogram() + #fill="white", position="dodge") +
  facet_wrap(~year)+
  xlab("duration (days)") 

dur_hist

p_duration <- ggplot(clean, aes(factor(month), fill = factor(year)))+
  geom_bar(position = "dodge") +
  #xlim(1,12)+
  facet_wrap(~tag.local.identifier)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

p_duration

```



# Banding locations

The banding locations were largely in Delaware Bay (included KIMBLESO, MOORES, FORTESCUE,  EASTPIT? East point?) and a location in New Jersey (Pesca) along with a location in Brazil near Porto Alegra (PEIXE).
This table shows the number of individuals by banding location and year.



```{r, echo = FALSE}
cleadid <- clean %>% 
  dplyr::select(individual.local.identifier, tag.local.identifier) %>%
  distinct()

dur <- left_join(dur_data, cleadid)

loc <- dur %>%
  dplyr::select(c(individual.local.identifier , dur_days))%>%
  mutate(dloc = stringr::str_sub(individual.local.identifier, 1, 4)) %>%
  mutate(dyr = stringr::str_sub(individual.local.identifier,13, 17)) %>%
  mutate(dyr = gsub("_", "", dyr))

dep_summary <- loc %>%
  dplyr::group_by(dloc, dyr)%>%
  dplyr::summarise( dloc_n = n())

#dep_summary

```

```{r, echo = FALSE}
knitr::kable(dep_summary)
```


# Number of active tags 

At the last data download, August 1st, all tags are inactive, in that those which have stopped transmitting. 

```{r , echo=FALSE}
#head(dur_data)

library(dplyr)
ddur <- dur_data %>%
  dplyr::mutate(max_month = month(max),
         max_year = year(max))
ddur <- ddur %>%
  dplyr::mutate(active = ifelse(max_month == 9 & max_year == 2023, "y", "n"))

active = ddur %>% filter(active == "y")
inactive = ddur %>% filter(active == "n")

```


Of the `r length(ddur$tag.local.identifier)` tags, `r length(active$tag.local.identifier)` were active, and `r length(inactive$tag.local.identifier)` were inactive.

```{r, echo = FALSE}

#sort(inactive$dur_days)

ggplot(ddur, aes(x = dur_days, y = as.factor(tag.local.identifier), colour=  active))+ 
  geom_point()


sort(ddur$v)


ddur <- ddur %>%
  mutate(dur_class = case_when(
    dur_days < 29 ~ 1,
    dur_days >29 & dur_days <60 ~2, 
    dur_days >60.1 & dur_days <100 ~3, 
    dur_days >100 ~4,
    .default = NA
 ))

```







# Geographic distributon of tags

```{r, echo=FALSE}
  
rf_sf <- sf::st_as_sf(clean, coords = c("location.long","location.lat"), crs = 4326, agr = "constant")

world <- ne_countries(scale = "medium", returnclass = "sf")

Americas <- world %>% dplyr::filter(region_un == "Americas")
#Americas <- world %>% dplyr::filter(continent == "North America")

# entire north America 
global <- ggplot(data = Americas) +
  geom_sf(color = "grey") +
  geom_sf(data = rf_sf, size = 1, colour = "dark blue") +
  #facet_wrap(~tag.local.identifier)+
  # geom_point(ru, aes(x = lng, y = lat), size = 4) +
  xlab("Longitude") + ylab("Latitude") +
  coord_sf(xlim = c(-130, -20), ylim = c(-50, 80), expand = FALSE)+
  theme_bw()+
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank())

global

# 
# jpeg(file.path(out.plots,"rose_density.jpg"), width = 30, height = 30,units = "cm", res = 210)
# # 2. Create the plot
# global 
```


```{r, echo = FALSE}

#head(alldat)


all <- ddur %>% 
  dplyr::select(tag.local.identifier , dur_days, year, dur_class)%>%
  left_join( cleadid)
  #rename("individual.local.identifier"  = individual.local.identifier.x )
   
arf_sf <- left_join(rf_sf, all) 

                        
# entire north America 
dd <- ggplot(data = Americas) +
  geom_sf(color = "grey") +
  geom_sf(data = arf_sf, size = 1, colour = "dark blue") +
  facet_wrap(~dur_class)+
  # geom_point(ru, aes(x = lng, y = lat), size = 4) +
  xlab("Longitude") + ylab("Latitude") +
  coord_sf(xlim = c(-130, -20), ylim = c(-50, 80), expand = FALSE)+
  theme_bw()+
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank())

#dd


```

# Plot based on duration class

 
```{r, echo = FALSE}
dd
```




# Detailed geographic locations vs sub populations

We can estimate the subpopulation level for those individuals with longer duration of tag timing?
Using the estimated ranges
- South East US
- Tierra Del Fuego
- Northern South America
- Western Gulf and western Pacific. 

As many of these individuals were banded at Delaware bay, this limits the subpops to SE_US, TDF, N_SA, while the WGWP group are less likely to be detected this far east. 


```{r, echo = FALSE}

min_lat <- arf_sf %>% 
  cbind(st_coordinates(arf_sf))%>%
  st_drop_geometry() %>%
  dplyr::select(individual.local.identifier, Y)%>%
  group_by(individual.local.identifier)%>%
  slice_min(Y) %>%
  mutate(pop_class = case_when(
     Y < -10 ~ "TDF",
     Y > -10.0001 & Y < 30 ~ "N_SA",
     #Y >  20.01 & Y < 30 ~ "SE_USA",
     Y > 30.1 ~ "USA",
      .default = NA
))


arf <- arf_sf %>%
  left_join(min_lat, by = "individual.local.identifier" )


# entire north America 
subpop <- ggplot(data = Americas) +
  geom_sf(color = "grey") +
  geom_sf(data = arf, size = 1, colour = "dark blue") +
  facet_wrap(~pop_class)+
  # geom_point(ru, aes(x = lng, y = lat), size = 4) +
  xlab("Longitude") + ylab("Latitude") +
  coord_sf(xlim = c(-130, -20), ylim = c(-50, 80), expand = FALSE)+
  theme_bw()+
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank())

subpop

```

# Breeding/extent locations: 


```{r}

a <- sort(unique(clean$tag.local.identifier ))

pvic <- c(213827, 213830, 213834 ,213836, 213838,  213841)

coats <- c(213831, 213835)

usa <- c(204359, 204362 ,213829, 213839, 213842, 224072, 224076, 224077 ,224078, 224079,224081,
         224086 ,224087,224091 ,224092, 224094 ,224095,  224096,224098, 224100, 224101 , 224102,
         224103)
         
cari <- c(204361, 224080, 224082, 224088,224097)

sth_am <- c(204351, 204352, 204357, 204364, 204369,  204370,  204375, 224073,  224075, 224083 ,
            224085, 224089, 224093, 224099, 233781)
            
hudson <- c(213828, 213832, 213833, 213837)
  
odd <- c(204371, 213840)



# breeding local 

breedsf <- arf_sf %>% 
  mutate(breed_class = case_when(
     tag.local.identifier %in%  usa ~ "usa",
     tag.local.identifier %in%  cari ~ "caribbean",
     tag.local.identifier %in% sth_am ~ "sth_america",
     tag.local.identifier %in%  hudson ~ "hudson",
     tag.local.identifier %in%  coats ~ "coats_is",
     tag.local.identifier %in%  pvic ~ "prince_vic",
     tag.local.identifier %in%  odd  ~ "unknown",
      .default = NA
))


# entire north America 
breed_loc <- ggplot(data = Americas) +
  geom_sf(color = "grey") +
  geom_sf(data = breedsf, size = 1, colour = "dark blue") +
  facet_wrap(~breed_class)+
  # geom_point(ru, aes(x = lng, y = lat), size = 4) +
  xlab("Longitude") + ylab("Latitude") +
  coord_sf(xlim = c(-130, -20), ylim = c(-50, 80), expand = FALSE)+
  theme_bw()+
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank())

breed_loc
     

```







Questions/Notes 

- many tags do not have tracks. All tracks have reference data, but many tags with reference data does not have tracks.  
- this data set includes the duplicate tracks for the spring migration dataset 
- Still to review the level of accuracy in more detail. 

Appendices: 

List of tags ids with no tracks or reference tags 

```{r, echo = FALSE}

norefnotrack = fullid %>%
  filter(is.na(individual.local.identifier))%>%
  #filter(animal.id == "") %>% 
  pull(tag.local.identifier)
  
norefnotrack


```


List of tags ids with no tracks

```{r, echo = FALSE}

refnotrack = fullid %>%
  filter(is.na(individual.local.identifier))%>%
  #filter(animal.id != "") %>% 
  pull(tag.local.identifier)
  
refnotrack
  


```
